Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2024-12-02 full_version/permutations.tex
minori

\begin{example}\todo{esempi di derivazione}
\end{example}

@@

\subsection{Colimits of derivations}
Given a DPO-rewriting system $(\X, \R)$,  we have already noted in \Cref{rem:func} that a derivation $\der{D}$  determines a diagram $\Delta(\der{D})$ in $\X$. We can then wonder if such a diagram has a colimit. Clearly if $\der{D}$ is the empty derivation $G$ then a colimit for $\Delta(\der{D})$ is simply the object $G$. More generally, we have the following result.

@@

@@

\subsection{Permutation equivalence}

\begin{definition}
	contenuto...
\end{definition}

\begin{theorem}[Characterization of $\equiv_p$]
	contenuto...
\end{theorem}
\begin{proof}
	contenuto...
\end{proof}

\section{Composing and restricting consistent permutations}
Now that we have established some initial facts about consistent permutations, we are going to embark in a, quite long, voyage to explore how to restrict them to suitable subderivations.

We can now turn our attention to the case in which a consistent permutation between two composite derivations is given. As a first step we need to prove some technicalities.

\begin{lemma}\label{prop:uniqu}
	Let $\sigma\colon [0,n]\to [0,n]$ be a consistent permutations between $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ and suppose that 
	\[(\der{D}, \alpha, \omega)=(\der{D}_1, \alpha_1, \omega_1) \cdot (\der{D}_2, \alpha_2, \omega_2) \qquad (\der{D}', \alpha', \omega')=(\der{D}'_1, \alpha'_1, \omega'_1) \cdot (\der{D}'_2, \alpha'_2, \omega'_2) \]

	 Let $l_1$ be the length of $\der{D}_1$, and suppose that there exists a consistent permutation $\tau:[0,l_1-1]\to [0, l_1-1]$.  If $q_1\colon \lpro \der{D}_1\rpro \to \lpro \der{D}\rpro$  and $q'_1\colon \lpro \der{D}'_1\rpro \to \lpro \der{D}'\rpro$ are the canonical arrows defined in \Cref{rem:dett}, then the following hold true:
	\begin{enumerate}
	\item  if $l_1=0$, then the following diagram commutes:
		\[\xymatrix{\lpro \der{D}_1\rpro \ar[r]^{\xi_\tau} \ar[d]_{q_1}& \lpro \der{D}'_1 \ar[d]^{q'_1}\rpro\\ \lpro \der{D}\rpro \ar[r]_{\xi_\sigma} & \lpro \der{D}'\rpro } \]
	%\[\xi_\sigma \circ q_1 \circ \iota_{1, G_{1,0}} = q'_1\circ \xi_{\tau} \circ  \iota_{1, G_{1,0}}\]
	\item if $l_1\neq 0$, so that $G_{1,0}=G_0$, then $\xi_\sigma \circ \iota_{G_0} = q'_1\circ \xi_{\tau} \circ  \iota_{1, G_{1,0}}$.
	\end{enumerate}

@@

We are now ready to prove the central result of this section: the uniqueness of a consistency permutation between two decorated derivations in a consuming rewriting system.

\begin{lemma}\label{lem:impo}
	Let $(\X,R)$ be a left-linear DPO-rewriting system and let also $(\der{D}_1, \alpha_1, \omega_1)$, $(\der{D}_2, \alpha_2, \omega_2)$ be two composable decorated derivations of length, respectively, $l_1$ and $l_2$. Suppose that $\sigma:[0, l_1+l_2-1]\to [0, l_1+l_2-1]$ is a consistent permutation between $(\der{D}_1, \alpha_1, \omega_1)\cdot (\der{D}_2, \alpha_2, \omega_2)$ and $(\der{D}'_1, \alpha'_1, \omega'_1)\cdot (\der{D}'_2, \alpha'_2, \omega'_2)$ for some other two composable decorated derivations $(\der{D}_1, \alpha_1, \omega_1)$ and $(\der{D}_2, \alpha_2, \omega_2)$.  Let, moreover, $\tau:[0,l_1-1]\to [0, l_1-1]$ be a consistent permutation between $(\der{D}_1, \alpha_1, \omega_1)$ and $(\der{D}'_1, \alpha'_1, \omega'_1)$. Suppose, finally, that  the set 
	\[D_{\sigma, \tau}:=\{i\in [0, l_1-1]\mid \sigma(i)\neq \tau(i)\}\]
	is non-empty and let $j$ be its minimum. If $r(\der{D}_1)$ is $\{\rho_i\}_{i=0}^{l_1-1}$ then the following hold true:
	\begin{enumerate}
		\item if $j=0$, then the rule $\rho_0$ is not consuming;
		\item if $j\neq 0$ then the rule $\rho_{j-1}$ is not consuming.
	\end{enumerate}
\end{lemma}
\begin{remark}\label{rem:minmax}It is worth to notice that the hypothesis $D_{\sigma, \tau} \neq \emptyset$ entails $l_1\neq 0$. Moreover, if $j\neq 0$, then $j-1$ is the maximum of $E_{\sigma, \tau}$.
\end{remark}
\begin{proof}
	\begin{enumerate}
		\item Let $k$ be $\sigma^{-1}(\tau(0))$ and notice that, since $\sigma(0)\neq \tau(0)$, then $0< k$.  By  the second point of  \Cref{prop:uniqu} we can consider the diagram
		\[\xymatrix@C=40pt{G_{0} \ar[drrr]_(.4){\iota_{G_0}}|(.675)\hole \ar[d]_{\iota_{1,G_{1,0}}}& L_0 \ar[r]_{m'_{\tau(0)}} \ar@/^.4cm/[rr]^{m_{k}} \ar[l]_{m_{0}} &G'_{\tau(0)} \ar[d]^(.45){\iota'_{G'_{\tau(0)}}} & G_{k} \ar[d]^{\iota_{G_k}}\\ \lpro\der{D}_1 \rpro \ar[r]_{\xi_{\tau}} &\lpro \der{D}'_1\rpro \ar[r]_{q'_1} & \lpro \der{D'}\rpro & \tpro{D} \ar[l]^{\xi_{\sigma}}}\]
		From \Cref{cor:ele}, we can conclude that there exists $c\colon L_0\to D_0$ such that $f_0\circ c=m_0$. We thus have the solid part of the commutative diagram below.
		\[\xymatrix{L_0 \ar@/^.3cm/[drr]^{\id{L_0}} \ar@{.>}[dr]_{t} \ar@/_.3cm/[ddr]_{c}\\ & K_0 \ar[d]_{k_0}\ar@{>->}[r]^{l_0}& L_0 \ar[d]^{m_0} \\& D_0 \ar@{>->}[r]_{f_0} & G_0} \]
		The internal square is an $\mathcal{M}$-pushout and thus a pullback, by \Cref{prop:pbpoad}, so that we have the existence of the dotted $t\colon L_0\to K_0$. Therefore $\id{L_0}=l_0\circ t$, proving that $l_0$ is an epimorphism. The thesis now follows from \Cref{cor:rego}. 
		\item  Let $k$ be $\sigma^{-1}(\tau(j-1))$ and notice that $\rho_{j-1}=\rho_k$. We have already noticed in \Cref{rem:minmax} that $j-1$ is the maximum of $E_{\sigma, \tau}$.  Hence, by the third point of \Cref{prop:uniqu} we have
		\[\xymatrix@C=40pt{G_{1,{j-1}} \ar[drrr]_(.4){\iota_{G_{j-1}}}|(.675)\hole \ar[d]_{\iota_{1,G_{1,j-1}}}& L_{j-1} \ar[r]_{m'_{\tau({j-1})}} \ar@/^.4cm/[rr]^{m_{k}} \ar[l]_{m_{{j-1}}} &G'_{\tau({j-1})} \ar[d]^{\iota'_{G'_{\tau({j-1})}}} & G_{k} \ar[d]^{\iota_{G_k}}\\ \lpro\der{D}_1 \rpro \ar[r]_{\xi_{\tau}} &\lpro \der{D}'_1\rpro \ar[r]_{q'_1} & \lpro \der{D'}\rpro & \tpro{D} \ar[l]^{\xi_{\sigma}}}\]
		Let $a$ be $\min(j-1, k)$, by \Cref{cor:ele},  there exists $c\colon L_{j-1}\to D_a$ such that $f_a\circ c=m_a$. As before this yields the solid part of the following diagram, in which the square is an $\mathcal{M}$-pushout.
		\[\xymatrix{L_{j-1} \ar@/^.3cm/[drr]^{\id{L_{j-1}}} \ar@{.>}[dr]_{t} \ar@/_.3cm/[ddr]_{c}\\ & K_{j-1} \ar[d]_{k_a}\ar@{>->}[r]^{l_{j-1}}& L_{j-1} \ar[d]^{m_a} \\& D_a \ar@{>->}[r]_{f_a} & G_a} \]

@@

On the one hand, computing we get:
		\begin{align*}
	q'_2\circ \psi \circ \iota_{2, G_{2,0}}\circ \alpha_2&=q \alpha_2\\&=\xi_\sigma \circ \iota_{G_{l_1+l_1}} \circ \omega_2\\&=\iota'_{G'_{l_1+l_2}}\circ \omega'_2\\&=q_2\circ \iota'_{2, G'_{2, l_2}}\circ \omega'_{2}
\end{align*}

		\begin{align*}
			q_2\circ \psi \circ \iota_{2, G_{2,l_2}}\circ \omega_2&=\xi_\sigma\circ q_2\circ \iota_{2, G_{2,l_2}}\circ \omega_2\\&=\xi_\sigma \circ \iota_{G_{l_1+l_1}} \circ \omega_2\\&=\iota'_{G'_{l_1+l_2}}\circ \omega'_2\\&=q_2\circ \iota'_{2, G'_{2, l_2}}\circ \omega'_{2}
		\end{align*}
	But $q_2$ is mono and then  $\psi \circ \iota_{2, G_{2,l_2}}\circ \omega_2=\iota'_{2, G'_{\varrho(i)}}\circ \omega'_{2}$ as wanted.

	On the other hand, by we have

	We proceed similarly for the back-matches:
	\begin{align*}
		q_2\circ \psi \circ \iota_{2, G_{2,i+1}}\circ h_{2,i}&=\xi_\sigma\circ q_2\circ \iota_{2, G_{2,i+1}}\circ h_{2,i}\\&=\xi_\sigma \circ \iota_{G_{i+l_1+1}} \circ h_{i+l_1}\\&=\iota'_{G'_{\sigma(i+l_1)+1}}\circ h'_{\sigma(i+l_1)}\\&=q_2\circ \iota'_{2, G'_{\varrho(i)+1}}\circ h'_{2, \varrho(i)}	
	\end{align*}

	For every

		Let now $i$ be an index different from $0$ and $\sigma^{-1}(l_1+1)$, then, using \Cref{rem:dett} and the previous point of this lemma we have:
		\begin{align*}	q_2\circ \psi \circ \iota_{2, G_{2,i}}\circ m_{2,i}&=\xi_\sigma\circ q_2\circ \iota_{2, G_{2,i}}\circ m_{2,i}\\&=\xi_\sigma \circ \iota_{G_{i+l_1}} \circ m_{i+l_1}\\&=\iota'_{G'_{\sigma(i+l_1)}}\circ m'_{\sigma(i+l_1)}\\&=q_2\circ \iota'_{2, G'_{\varrho(i)}}\circ m'_{2, \varrho(i)}
		\end{align*}

	Using the fact that $q_2$ belongs to $\mathcal{M}$ we conclude that 
	\[\psi \circ \iota_{2, G_{2,i+1}}\circ h_{2,i}=\iota'_{2, G'_{\varrho(i)+1}}\circ h'_{2, \varrho(i)}\]

		We are left with two cases. First of all, we can notice thatIn this case we have:
		\[G_{i}=\pi(G_{2,0}) \quad G'_{i}=\pi(G'_{2,0}) \quad m_{l_1}= \alpha^{-1}_2 \circ m_{2,0} \quad m'_{l_1}= (\alpha^{-1}_2)^{-1} \circ m'_{2,0} \]

		Let's compute:
		\[q_2\circ \psi \circ \iota_{2}\]

		Consistency now follows at once. \qedhere 
	\end{enumerate}
\end{proof}

Finally, we  end this section showing that consistent permutations between composable abstract derivations yield a consistent permutation between the composites.

	Given a consistent permutation $\sigma\colon [0, \lgh(\der{D}_1)-1]\to [0, \lgh(\der{D}_1)-1]$ between  $(\der{D}_1, \alpha_1, \omega_1)$ and $(\der{D}'_1, \alpha'_1, \omega'_1)$ and another $\tau\colon [0, \lgh(\der{D}_2)-1]\to [0, \lgh(\der{D}_2)-1]$ between $(\der{D}_2, \alpha_2, \omega_2)$ and $(\der{D}'_2, \alpha'_2, \omega'_2)$, then:
	\[\sigma+\tau\colon[0, \lgh(\der{D})-1]\to[0, \lgh(\der{D})-1] \quad i \mapsto \begin{cases}
		\sigma(i) & i < \lgh(\der{D}_1)\\
		\lgh(\der{D}_1)+\tau(i-\lgh(\der{D}_1)) &   i\geq \lgh(\der{D}_1) 
	\end{cases}\]
	is a consistent permutation between $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$.
\end{lemma}

\begin{proof}

Given $X_1\in \Delta(\der{D}_1)$, $X_2\in \Delta(\der{D}_2)$, $X'_1\in \Delta(\der{D}'_1)$, $X'_2\in \Delta(\der{D}'_2)$, we will denote their coprojections into, respectively, $\lpro \der{D}_1\rpro$, $\lpro \der{D}'_2\rpro$, $\lpro \der{D}'_1\rpro$, $\lpro \der{D}'_2\rpro$ by
$\iota_{1, X_1}\colon X_1\to \lpro\der{D}_1\rpro$, $\iota_{1, X_2}\colon X_2\to \lpro\der{D}_1\rpro$, $\iota'_{1, X'_1}\colon X'_1\to \lpro\der{D}'_1\rpro$ and $\iota'_{2, X'_2}\colon X'_2\to \lpro\der{D}_1\rpro$. Similarly, given $X\in \Delta(\der{D})$ and $X'\in \Delta(\der{D}')$ we will use $\iota_X\colon X\to \tpro{D}$ and $\iota'_{X'}\colon X'\to \lpro \der{D}' \rpro$ for the coprojections.

We can sum up the results of this section recasting them in the language of permutation equivalence.

\begin{lemma}
	contenuto...
\end{lemma} 
\begin{proof}
	contenuto...
\end{proof}

>>>>

@@

\subsection{Colimits of derivations}\label{subsec:col}
Given a DPO-rewriting system $(\X, \R)$,  we have already noted in \Cref{rem:func} that a derivation $\der{D}$  determines a diagram $\Delta(\der{D})$ in $\X$. We can then wonder if such a diagram has a colimit. Clearly if $\der{D}$ is the empty derivation $G$ then a colimit for $\Delta(\der{D})$ is simply the object $G$. More generally, we have the following result.

@@

We want now to relate the colimit of a composite decorated derivation with the colimits of its components. The following two lemmas will be pivotal to develop the subsequent sections.

@@

We have already proved in \Cref{rem:abscons} that the identity defines a consistent permutation between any two representatives of an equivalence class $[\der{D}, \alpha, \omega]_a$. This, together with \Cref{rem:comp}, allow us to immediately derive the following fact.

\begin{corollary}\label{prop:abs}
Let $(\der{D}_1, \alpha_1, \omega_1)$ and $(\der{D}_2, \alpha_2, \omega_2)$ be two decorated derivations and suppose that $\sigma$ is a consistent permutation between them. If  $(\der{D}'_1, \alpha'_1, \omega'_1)$ and $(\der{D}'_2, \alpha'_2, \omega'_2)$ are such that
\[ (\der{D}_1, \alpha_1, \omega_1)\equiv^a (\der{D}'_1, \alpha'_1, \omega'_1) \qquad (\der{D}_2, \alpha_2, \omega_2)\equiv^a (\der{D}'_2, \alpha'_2, \omega'_2) \] 
then $\sigma$ is a consistent permutation between $(\der{D}'_1, \alpha'_1, \omega'_1)$ and $(\der{D}'_2, \alpha'_2, \omega'_2)$ too.
\end{corollary}

\begin{proof} Let $l_1$ and $l_2$ be the lengths of $\der{D}_1$ and $\der{D}_2$. By \Cref{rem:abscons} we know that  $\id{[0, l_1-1]}$ is a consistent permutation between  $(\der{D}_1, \alpha_1, \omega_1)$ and $(\der{D}'_1, \alpha'_1, \omega'_1)$  and $\id{[0, l_2-1]}$ one between   $(\der{D}_2, \alpha_2, \omega_2)$  and $(\der{D}'_2, \alpha'_2, \omega'_2)$. Using \Cref{rem:comp} we get the thesis at once.
\end{proof}

The previous result suggests that the existence of a consistent permutations induces an equivalence relation on the arrows of $\dpi$. This is formalized by the next definition.

\begin{definition}Let $(\X, \R)$ be a left-linear DPO rewriting system. We say that two arrows
	$[\der{D}, \alpha, \omega]_a, [\der{D}', \alpha', \omega']_a\colon [G]\rightrightarrows [H]$ in $\dpi$ are \emph{permutation equivalent} if there exists a consistent permutation between $(\der{D}, \alpha, \omega)$  and $(\der{D}', \alpha', \omega')$.

	We will use the notation $[\der{D}, \alpha, \omega]_a, \equiv^p [\der{D}', \alpha', \omega']_a$  to denote permutation equivalence, while we will use $[\der{D}, \alpha, \omega]_a, \equiv_\sigma^p [\der{D}', \alpha', \omega']_a$ to point out a witness $\sigma$ to this relation.
\end{definition}

\begin{remark}
	The relation $\equiv^p$ is an equivalence: \Cref{prop:abs} guarantees that it is well-defined, \Cref{rem:abscons} guarantees reflexivity, while symmetry and transitivity follow from, respectively \Cref{rem:inversa} and \Cref{rem:comp}.
\end{remark}

\begin{notation}
	We will denote by $[\der{D}, \alpha, \omega]_p$ the equivalence class of $[\der{D}, \alpha, \omega]_a$ with respect to the permutation equivalence.
\end{notation}

\section{Composing and restricting consistent permutations}

We want now to exploit the results proved at the end of \Cref{subsec:col} to study the behaviour of permutation equivalence with respect to composition of decorated derivations. As a first step we need to establish some technical points.

\begin{lemma}\label{prop:uniqu}
	Let $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$ be two decorated derivations such that $[\der{D}, \alpha, \omega]_a\equiv^p_\sigma [\der{D}', \alpha', \omega']_a $. Suppose that 
	\[(\der{D}, \alpha, \omega)=(\der{D}_1, \alpha_1, \omega_1) \cdot (\der{D}_2, \alpha_2, \omega_2) \qquad (\der{D}', \alpha', \omega')=(\der{D}'_1, \alpha'_1, \omega'_1) \cdot (\der{D}'_2, \alpha'_2, \omega'_2) \]

	 Let $l_1$ and $l_2$ be the lengths of, respectively, $\der{D}_1$ and $\der{D}_2$ and suppose that  $[\der{D}_1, \alpha_1, \omega_1]_a\equiv^p_\tau [\der{D}'_1, \alpha'_1, \omega'_1]_a$ for some consistent permutation $\tau\colon [0, l_1-1]\to [0,l_1-1]$.  If $q_1\colon \lpro \der{D}_1\rpro \to \lpro \der{D}\rpro$  and $q'_1\colon \lpro \der{D}'_1\rpro \to \lpro \der{D}'\rpro$ are the canonical arrows defined in \Cref{rem:dett}, then the following hold true:
	\begin{enumerate}
	\item  if $l_1=0$, then the following diagram commutes:
		\[\xymatrix{\lpro \der{D}_1\rpro \ar[r]^{\xi_\tau} \ar[d]_{q_1}& \lpro \der{D}'_1 \ar[d]^{q'_1}\rpro\\ \lpro \der{D}\rpro \ar[r]_{\xi_\sigma} & \lpro \der{D}'\rpro } \]
	%\[\xi_\sigma \circ q_1 \circ \iota_{1, G_{1,0}} = q'_1\circ \xi_{\tau} \circ  \iota_{1, G_{1,0}}\]
	\item if $l_1\neq 0$, so that $G_{1,0}=G_0$, then $\xi_\sigma \circ \iota_{G_0} = q'_1\circ \xi_{\tau} \circ  \iota_{1, G_{1,0}}$.
	\end{enumerate}

@@

We want now to show that permutation equivalence is a congruence with respect to composition in $\dpi$. This, in turn, will allow us to build  a category in which arrows are are exactly the equivalence classes of permutation equivalent abstract decorated decorations.

\begin{remark}\label{rem:sum}
	Let $n$ and $m$ be two natural numbers and $\sigma\colon [0, n-1]\to [0,n-1]$ and $\tau\colon [0, m-1]\to [0,m-1]$ be two permutations. Then they induce a permutation $[0, n+m-1]\to [0, n+m-1]$. To see this, it is enough to define
		\[\sigma+\tau\colon[0, n+m-1]\to[0, n+m-1] \quad i \mapsto \begin{cases}
		\sigma(i) & i < n\\
		\tau(i-n) + n &  n\leq i 
	\end{cases}\]
\end{remark}

Suppose that $[\der{D}_1, \alpha_1, \omega_1]_a \equiv^p_\sigma [\der{D}'_1, \alpha'_1, \omega'_1]_a$ and $[\der{D}_2, \alpha_2, \omega_2]_a \equiv^p_\tau [\der{D}'_2, \alpha'_2, \omega'_2]_a$, then:
\[[\der{D}, \alpha, \omega]_a\equiv^p_{\sigma+\tau} [\der{D}', \alpha', \omega']_a\]
\end{lemma}
\begin{proof}
By \Cref{prop:abs}	it is enough to show that $\sigma+\tau$	is a consistent permutation between $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$.

We need to fix some notation.	Given $X_1\in \Delta(\der{D}_1)$, $X_2\in \Delta(\der{D}_2)$, $X'_1\in \Delta(\der{D}'_1)$, $X'_2\in \Delta(\der{D}'_2)$, we will denote their coprojections into, respectively, $\lpro \der{D}_1\rpro$, $\lpro \der{D}'_2\rpro$, $\lpro \der{D}'_1\rpro$, $\lpro \der{D}'_2\rpro$ by
	$\iota_{1, X_1}\colon X_1\to \lpro\der{D}_1\rpro$, $\iota_{1, X_2}\colon X_2\to \lpro\der{D}_1\rpro$, $\iota'_{1, X'_1}\colon X'_1\to \lpro\der{D}'_1\rpro$ and $\iota'_{2, X'_2}\colon X'_2\to \lpro\der{D}_1\rpro$. Similarly, given $X\in \Delta(\der{D})$ and $X'\in \Delta(\der{D}')$ we will use $\iota_X\colon X\to \tpro{D}$ and $\iota'_{X'}\colon X'\to \lpro \der{D}' \rpro$ for the coprojections.

\begin{definition}
Let $(\X, \R)$ be a left-linear DPO rewriting system. The category $\ppi$ is defined as follows: 
	\begin{itemize}
	\item objects are isomorphism classes of objects of $\X$;
	\item an arrow $[G]\to [H]$ is an equivalence class $[\der{D}, \alpha, \omega]_p$ of abstract decorated derivations between $[G]$ and $[H]$;
	\item given their composition is defined as
	\[[\der{D}_2, \alpha_2, \omega_2]_p \circ [\der{D}_1, \alpha_1, \omega_1]_p:=[(\der{D}_1, \alpha_1, \omega_1) \cdot (\der{D}_2, \alpha_2, \omega_2)]_p\] 
	\item the identity on $[G]$ is $[G, \alpha, \alpha]_p$, where $\alpha$ is any isomorphism $\pi(G)\to G$.	\end{itemize}
\end{definition}
\begin{remark}By \Cref{rem:empty2}
\end{remark}

We are now going to proceed towards the central result of this section: the uniqueness of a consistency permutation between two decorated derivations in a consuming rewriting system.

\begin{lemma}\label{lem:impo}
Let  $(\der{D}_1, \alpha_1, \omega_1)$, $(\der{D}_2, \alpha_2, \omega_2)$ be two composable decorated derivations of length, respectively, $l_1$ and $l_2$  in a left-linear DPO-rewriting system $(\X,R)$. Let also  $(\der{D}'_1, \alpha'_1, \omega'_1)$ and $(\der{D}'_2, \alpha'_2, \omega'_2)$ be other two composable decorated derivations such that:
	\begin{align*}(\der{D}_1, \alpha_1, \omega_1)&\equiv^p_\tau(\der{D}'_1, \alpha'_1, \omega'_1)\\
	(\der{D}_1, \alpha_1, \omega_1)\cdot (\der{D}_2, \alpha_2, \omega_2)&\equiv^p_\sigma (\der{D}'_1, \alpha'_1, \omega'_1)\cdot (\der{D}'_2, \alpha'_2, \omega'_2)
	\end{align*}

Suppose, moreover, that the set
	\[D_{\sigma, \tau}:=\{i\in [0, l_1-1]\mid \sigma(i)\neq \tau(i)\}\]
	is non-empty and let $j$ be its minimum. If $r(\der{D}_1)$ is $\{\rho_i\}_{i=0}^{l_1-1}$ then the following hold true:
	\begin{enumerate}
		\item if $j=0$, then the rule $\rho_0$ is not consuming;
		\item if $j\neq 0$ then the rule $\rho_{j-1}$ is not consuming.
	\end{enumerate}
\end{lemma}
\begin{remark}\label{rem:minmax}It is worth to notice that the hypothesis $D_{\sigma, \tau} \neq \emptyset$ entails $l_1\neq 0$. Moreover, if $j\neq 0$, then $j-1$ is the maximum of $E_{\sigma, \tau}$.
\end{remark}
\begin{proof}
	\begin{enumerate}
		\item Let $k$ be $\sigma^{-1}(\tau(0))$ and notice that, since $\sigma(0)\neq \tau(0)$, then $0< k$.  By  the second point of  \Cref{prop:uniqu} we can consider the diagram
		\[\xymatrix@C=40pt{G_{0} \ar[drrr]_(.4){\iota_{G_0}}|(.675)\hole \ar[d]_{\iota_{1,G_{1,0}}}& L_0 \ar[r]_{m'_{\tau(0)}} \ar@/^.4cm/[rr]^{m_{k}} \ar[l]_{m_{0}} &G'_{\tau(0)} \ar[d]^(.45){\iota'_{G'_{\tau(0)}}} & G_{k} \ar[d]^{\iota_{G_k}}\\ \lpro\der{D}_1 \rpro \ar[r]_{\xi_{\tau}} &\lpro \der{D}'_1\rpro \ar[r]_{q'_1} & \lpro \der{D'}\rpro & \tpro{D} \ar[l]^{\xi_{\sigma}}}\]
		From \Cref{cor:ele}, we can conclude that there exists $c\colon L_0\to D_0$ such that $f_0\circ c=m_0$. We thus have the solid part of the commutative diagram below.
		\[\xymatrix{L_0 \ar@/^.3cm/[drr]^{\id{L_0}} \ar@{.>}[dr]_{t} \ar@/_.3cm/[ddr]_{c}\\ & K_0 \ar[d]_{k_0}\ar@{>->}[r]^{l_0}& L_0 \ar[d]^{m_0} \\& D_0 \ar@{>->}[r]_{f_0} & G_0} \]
		The internal square is an $\mathcal{M}$-pushout and thus a pullback, by \Cref{prop:pbpoad}, so that we have the existence of the dotted $t\colon L_0\to K_0$. Therefore $\id{L_0}=l_0\circ t$, proving that $l_0$ is an epimorphism. The thesis now follows from \Cref{cor:rego}. 
		\item  Let $k$ be $\sigma^{-1}(\tau(j-1))$ and notice that $\rho_{j-1}=\rho_k$. We have already noticed in \Cref{rem:minmax} that $j-1$ is the maximum of $E_{\sigma, \tau}$.  Hence, by the third point of \Cref{prop:uniqu} we have
		\[\xymatrix@C=40pt{G_{1,{j-1}} \ar[drrr]_(.4){\iota_{G_{j-1}}}|(.675)\hole \ar[d]_{\iota_{1,G_{1,j-1}}}& L_{j-1} \ar[r]_{m'_{\tau({j-1})}} \ar@/^.4cm/[rr]^{m_{k}} \ar[l]_{m_{{j-1}}} &G'_{\tau({j-1})} \ar[d]^{\iota'_{G'_{\tau({j-1})}}} & G_{k} \ar[d]^{\iota_{G_k}}\\ \lpro\der{D}_1 \rpro \ar[r]_{\xi_{\tau}} &\lpro \der{D}'_1\rpro \ar[r]_{q'_1} & \lpro \der{D'}\rpro & \tpro{D} \ar[l]^{\xi_{\sigma}}}\]
		Let $a$ be $\min(j-1, k)$, by \Cref{cor:ele},  there exists $c\colon L_{j-1}\to D_a$ such that $f_a\circ c=m_a$. As before this yields the solid part of the following diagram, in which the square is an $\mathcal{M}$-pushout.
		\[\xymatrix{L_{j-1} \ar@/^.3cm/[drr]^{\id{L_{j-1}}} \ar@{.>}[dr]_{t} \ar@/_.3cm/[ddr]_{c}\\ & K_{j-1} \ar[d]_{k_a}\ar@{>->}[r]^{l_{j-1}}& L_{j-1} \ar[d]^{m_a} \\& D_a \ar@{>->}[r]_{f_a} & G_a} \]

@@

On the one hand, computing we get:
		\begin{align*}
\psi \circ \iota_{2, G_{2,0}}\circ \alpha_2&=  \zeta_0\circ \alpha_2\\&=  \iota'_{2, G'_{2,0}}\circ \alpha'_2\circ \alpha^{-1}_2\circ \alpha_2\\&=\iota'_{2, G'_{2,0}}\circ \alpha'_2
\end{align*}
	On the other hand, by we have
		\begin{align*}
			q'_2\circ \psi \circ \iota_{2, G_{2,l_2}}\circ \omega_2&=\xi_\sigma\circ q_2\circ \iota_{2, G_{2,l_2}}\circ \omega_2\\&=\xi_\sigma \circ \iota_{G_{l_1+l_1}} \circ \omega_2\\&=\iota'_{G'_{l_1+l_2}}\circ \omega'_2\\&=q'_2\circ \iota'_{2, G'_{2, l_2}}\circ \omega'_{2}
		\end{align*}
	But $q'_2$ is mono and then  $\psi \circ \iota_{2, G_{2,l_2}}\circ \omega_2=\iota'_{2, G'_{l_2}}\circ \omega'_{2}$ as wanted.

	We proceed similarly for the back-matches:
	\begin{align*}
		q'_2\circ \psi \circ \iota_{2, G_{2,i+1}}\circ h_{2,i}&=\xi_\sigma\circ q_2\circ \iota_{2, G_{2,i+1}}\circ h_{2,i}\\&=\xi_\sigma \circ \iota_{G_{i+l_1+1}} \circ h_{i+l_1}\\&=\iota'_{G'_{\sigma(i+l_1)+1}}\circ h'_{\sigma(i+l_1)}\\&=q'_2\circ \iota'_{2, G'_{\varrho(i)+1}}\circ h'_{2, \varrho(i)}	
	\end{align*}
	Again we conclude since $q'_2$ is a monomorphism.

	We turn now to consistency of $\varrho$ with respect to the matches. If $i$ is an index different from $0$ then we have:
		\begin{align*}	q_2\circ \psi \circ \iota_{2, G_{2,i}}\circ m_{2,i}&=\xi_\sigma\circ q_2\circ \iota_{2, G_{2,i}}\circ m_{2,i}\\&=\xi_\sigma \circ \iota_{G_{i+l_1}} \circ m_{i+l_1}\\&=\iota'_{G'_{\sigma(i+l_1)}}\circ m'_{\sigma(i+l_1)} \\&= 
			\begin{cases}
			\iota'_{G'_{l_1}}\circ (\alpha'_2)^{-1} \circ m'_{2, 0}  &\sigma(i+l_1) = l_1\\
				\iota'_{G'_{\sigma(i+l_1)}}\circ m'_{2,\varrho(i)} & \sigma(i+l_1)\neq l_1
			\end{cases}
			\\&=\begin{cases}
			q'_2\circ \iota_{2, G'_{2,0}} \circ \alpha'_2\circ  (\alpha'_1)^{-1} \circ m'_{2, 0}& \varrho(i)=0\\
			q'_2\circ \iota'_{G'_{2, \varrho(i)}} \circ m'_{2,\varrho(i)} & \varrho(i)\neq 0
			\end{cases}\\&=q_2\circ \iota'_{2, G'_{\varrho(i)}}\circ m'_{2, \varrho(i)}
		\end{align*}
		and the thesis follows since $q'_2$ belongs to $\mathcal{M}$.

	 We are thus left with the case $i=0$. The argument is more or less the same as before:
	 \begin{align*}
	 	q_2\circ \psi \circ \iota_{2, G_{2,0}}\circ m_{2,0}&=\xi_\sigma \circ q_2\circ \iota_{2, G_{2,0}}\circ m_{2,0}\\&=\xi_\sigma \circ \iota_{G_{l_1}}\circ \alpha^{-1}_2\circ m_{2,0}\\&=\xi_\sigma \circ \iota_{ G_{l_1}}\circ m_{l_1}\\&=\iota'_{G_{\sigma(l_1)}}\circ m'_{\sigma(l_1)}\\&=\begin{cases}
	 		\iota'_{G'_{l_1}}\circ (\alpha'_2)^{-1} \circ m'_{2, 0}  &\sigma(l_1) = l_1\\
	 		\iota'_{G'_{\sigma(l_1)}}\circ m'_{2,\varrho(0)} & \sigma(l_1)\neq l_1
	 	\end{cases}
	 	\\&=\begin{cases}
	 		q'_2\circ \iota_{2, G'_{2,0}} \circ \alpha'_2\circ  (\alpha'_1)^{-1} \circ m'_{2, 0}& \varrho(i)=0\\
	 		q'_2\circ \iota'_{G'_{2, \varrho(i)}} \circ m'_{2,\varrho(i)} & \varrho(i)\neq 0
	 	\end{cases}\\&=q_2\circ \iota'_{2, G'_{\varrho(i)}}\circ m'_{2, \varrho(i)}
	 \end{align*}
	 But $q_2$ is a mono and so we can conclude.	  \qedhere 
	\end{enumerate}
\end{proof}

\begin{corollary}\label{cor:epi}
	contenuto...
\end{corollary}
